---
title: "Writing Functions"
author: "Derek Lamb"
date: "`r Sys.Date()`"
output: github_document
---

```{r load packages}
library(tidyverse)
library(rvest)
library(knitr)

# Set default figure options
opts_chunk$set(
  fig.width = 6,
  out.width = "90%"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Set seed for reproducibility.

```{r set seed}
set.seed(12345)
```


## Z score function
Z scores subtract the mean and devide by the sd.

```{r z score function}
x_vec <- rnorm(20, mean = 5, sd = 0.3)
```

```{r}
(x_vec - mean(x_vec))/sd(x_vec)
```

Write a function to do this!
```{r}
z_score = function(x) {
  
  if (!is.numeric(x)){
    stop("Argument should be numbers")
  }
  else if (length(x) < 2){
    stop("You need at least two numbers to get z scores")
  }
  
  z = (x - mean(x)) / sd(x)
  return(z)
}
```

Check that this works.

```{r}
z_score(x = x_vec)
```

Keep checking

```{r, error = TRUE}
z_score(x = 3)

z_score(c("My", "name", "is", "Jeff"))

z_score(c(TRUE, TRUE, FALSE, TRUE))

z_score(iris)
```

### Multiple outputs

Write a function that returns the mean and sd from a sample.
```{r}
mean_and_sd = function(x) {

  if (!is.numeric(x)){
    stop("Argument should be numbers")
  }
  else if (length(x) < 2){
    stop("You need at least two numbers to get standard deviations")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  tibble(
    mean = mean_x,
    sd = sd_x
  )
}
```

Double check this works
```{r}
mean_and_sd(x_vec)
```


### Start geting means and sds

```{r}

x_vec = rnorm(n = 30, mean = 5, sd = 0.5)

  tibble(
    mean = mean(x_vec),
    sd = sd(x_vec)
  )

```

Let's write a function that uses `n`, a true mean, and a true sd as inputs.

```{r}
sim_mean_sd = function(n_obs, mu = 5, sigma = 1){

  x_vec = rnorm(n = n_obs, mean = mu, sd = sigma)

  tibble(
    mean = mean(x_vec),
    sd = sd(x_vec)
  )
  
}
```

```{r}
sim_mean_sd(n_obs = 50)

sim_mean_sd(mu = 12,n_obs = 24,4)
```


### Learning check (LotR)
```{r}
lotr_reader = function(cells, name, path = "data/LotR_Words.xlsx") {
  
  if (!is.character(cells)) {
    stop("Cells should be a character")
  }
  
  df = readxl::read_excel(path, range = cells) |> 
    mutate(movie = name,
           Race = str_to_lower(Race)) |> 
    janitor::clean_names() |> 
    pivot_longer(
      female:male,
      names_to = "sex",
      values_to = "word_count"
    ) |> 
    select(movie, everything())
  
  return(df)
}
```

Check if this works
```{r}
df_lotr = bind_rows(
  lotr_reader("B3:D6", "fellowship_ring"),
  lotr_reader("F3:H6", "two_towers"),
  lotr_reader("J3:L6", "return_king")
)
```


```{r}

nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)


data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

```{r}
nsduh_reader = function(n_table, outcome_name, url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm") {
  
  df = 
  read_html(url) |> 
  html_table() |> 
  nth(n_table) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent),
    outcome = outcome_name) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  return(df)
}

```

```{r}

df_marj = nsduh_reader(1, "marijuana")
```


